---
title: "BaslineTable"
author: "Friederike Gutmann"
date: "`r Sys.Date()`"
output: html_document
---

# Preparation

## Load packages

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
source("Packages_Metabolites.R", local = knitr::knit_global())
```


## Data

*This part needs to be adjusted for each project!*

### SampleIDs

Load **sampleID list** to connect metabolomics- with metadata. The column names in the loaded file should be the same as the corresponding ones in the peak and metadata files, otherwise the merge will not work. Additionally, the peak and the meta data will be adjusted according to this file, so they only include the samples listed in that file (and vice versa). Add the column name that is connecting the sampleID file with the metadata file in `meta.merge`.

```{r Connection data, echo = FALSE}
PerProtocol_samples <- read.csv(here("Samples", "SampleIDs_PerProtocol_V1V3.csv"), header = TRUE) %>%
  select(-Sample.Number) %>%
  filter(Visit == "V0")

dropout_samples <- read.csv(here("Samples", "SampleIDs_DropOuts_V0_wTubeID.csv"), header = TRUE) %>%
  select(-Sample_Number, -Tube_ID)

```


### Metadata
Load and inspect the **metadata**.

```{r Metadata start, echo = TRUE}
disease_duration <- read.csv(here("Metadata", "Disease_duration.csv"), header = TRUE, dec = ",") %>%
  rename("PatientID" = "IDs")

ms_meds <- read.csv(here("Metadata", "NAMS_Medication_pervisit.csv"), header = T, dec = ",") %>%
  filter(grepl("v0|v3", redcap_event_name)) %>%
  rename("PatientID" = "record_id") %>%
  filter(if_all(everything(), ~ . != "")) %>%
  mutate(MS.Med = str_replace_all(MS.Med, "no med", "No_DMT")) %>%
  mutate(MS.Med = str_replace_all(MS.Med, "No med", "No_DMT")) %>%
  mutate(MS.Med = str_replace_all(MS.Med, "Privigen", "No_DMT")) %>% #Privigen is not a DMT
  mutate(MS.Med = str_replace_all(MS.Med, "Avonex", "Plegridy")) %>% #both infb1a
  mutate(MS.Med = str_replace_all(MS.Med, "Gylenia", "Fingolimod"))


ms_meds_V0_dropout <- ms_meds %>% 
  filter(grepl("v0", redcap_event_name)) %>%
  filter(PatientID %in% dropout_samples$PatientID) %>%
  distinct(PatientID, MS.Med, .keep_all = TRUE) %>%
  mutate(present = TRUE) %>%
  pivot_wider(
    id_cols = PatientID,
    names_from = MS.Med,
    values_from = present,
    values_fill = FALSE
  )

ms_meds_V0_pp <- ms_meds %>% 
  filter(grepl("v0", redcap_event_name)) %>%
  filter(PatientID %in% PerProtocol_samples$PatientID) %>%
  distinct(PatientID, MS.Med, .keep_all = TRUE) %>%
  mutate(present = TRUE) %>%
  pivot_wider(
    id_cols = PatientID,
    names_from = MS.Med,
    values_from = present,
    values_fill = FALSE
  ) 

#ms_meds_filtered <- ms_meds %>%
  #mutate(MS.Med = str_replace_all(MS.Med, "L04AA36", "Ocrelizumab")) %>%
  #mutate(MS.Med = str_replace_all(MS.Med, "L04AA23", "Natalizumab")) %>%
  #mutate(MS.Med = str_replace_all(MS.Med, "L04AA27", "Fingolimod")) %>%
  #mutate(MS.Med = str_replace_all(MS.Med, "L04AA31", "Teriflunomide")) %>%
  #mutate(MS.Med = str_replace_all(MS.Med, "L04AA34", "Alemtuzumab")) %>%
  #mutate(MS.Med = str_replace_all(MS.Med, "L04AC01", "Daclizumab")) %>%
  #mutate(MS.Med = str_replace_all(MS.Med, "L04AX07", "DMF")) %>%
  #mutate(MS.Med = str_replace_all(MS.Med, "L03AB07", "Ifba")) %>%
  #mutate(MS.Med = str_replace_all(MS.Med, "L03AB08", "Ifbb")) %>%
  #mutate(MS.Med = str_replace_all(MS.Med, "L03AX13", "Glaace")) %>%
  #mutate(MS.Med = str_replace_all(MS.Med, "L01XC02", "Rituximab")) 


#write_csv(ms_meds_V0_pp, here("Metadata", "MS_meds_clean.csv"))


metadata <- read.csv(here("Metadata", "NAMS_Data_FGcurated.csv"), header = TRUE, dec = ",") %>%
  merge(disease_duration, by = "PatientID") %>%
  mutate(across(c(PatientID, Group, scrn_ethnicity.factor, info_gender.factor, lesion_load.v0_arm_1, lesion_load.v3_arm_1, lesion_load.v5_arm_1, dropout_prot_reason), ~ as.factor(.))) %>%
  mutate_if(is.character, as.numeric) %>%
  rename_with(~ gsub("_arm_1", "", .x)) %>%
  rename_with(~ gsub("_V", ".v", .x)) %>%
  select(-matches("_V5$|.v5"), -matches("_V3$|.v3")) %>%
  mutate(waist_hip_ratio.v0 = mes_waist_circum.v0/mes_hip_circum_cm.v0)

columns_with_inf <- sapply(metadata, function(x) any(is.infinite(x)))
inf_columns <- names(columns_with_inf[columns_with_inf])
metadata <- metadata %>% select(-inf_columns)

#write.csv(metadata, here("Metadata", "NAMS_Metadata_for_analysis_allDiffs_wDiv_gmmsfpkm.csv"))
```


### Metadata for each subgroup of sample (per protocol and dropouts)

```{r Metadata subgroups, echo = TRUE}
metadata_pp <- metadata %>%
  merge(ms_meds_V0_pp, by = "PatientID") %>%
  filter(PatientID %in% PerProtocol_samples$PatientID)

metadata_do <- metadata %>%
  merge(ms_meds_V0_dropout, by = "PatientID") %>%
  filter(PatientID %in% dropout_samples$PatientID)
```

# Baseline Tabelle

## Age
```{r Age Stats, echo = TRUE}
stats_func <- function(df) {
  group_stats_func <- df %>%
    group_by(Group) %>%
    summarise(
      n = n(),
      mean_age = round(mean(visit_age.v0, na.rm = TRUE)),
      sd_age = round(sd(visit_age.v0, na.rm = TRUE))
    )
  
  overall_stats_func <- df %>%
    summarise(
      group = NULL,
      n = n(),
      mean_age = round(mean(visit_age.v0, na.rm = TRUE)),
      sd_age = round(sd(visit_age.v0, na.rm = TRUE))
    )
  
  return(bind_rows(group_stats_func, overall_stats_func))
  
}

stats_func(metadata_pp)
stats_func(metadata_do)
kruskal.test(visit_age.v0 ~ Group, data = metadata_pp)
kruskal.test(visit_age.v0 ~ Group, data = metadata_do)
```


## Ethnicity
```{r Ethnicity Stats, echo = TRUE}
stats_ethnicity <- function(df) {
  group_stats_func <- df %>%
    group_by(Group) %>%
    summarise(
      n = n(),
      caucasian_n = sum(scrn_ethnicity.factor == "kaukasisch/weiß", na.rm = TRUE),
      caucasian = sum(scrn_ethnicity.factor == "kaukasisch/weiß", na.rm = TRUE)/n())
  
  overall_stats_func <- df %>%
    summarise(
      group = NULL,
      n = n(),
      caucasian_n = sum(scrn_ethnicity.factor == "kaukasisch/weiß", na.rm = TRUE),
      caucasian = sum(scrn_ethnicity.factor == "kaukasisch/weiß", na.rm = TRUE)/n())
  
  return(bind_rows(group_stats_func, overall_stats_func))
  
}

stats_ethnicity(metadata_pp)
stats_ethnicity(metadata_do)

chisq.test(metadata_pp$Group, metadata_pp$scrn_ethnicity.factor)
chisq.test(metadata_do$Group, metadata_do$scrn_ethnicity.factor)

```



## Gender
```{r Ethnicity Stats, echo = TRUE}
stats_ethnicity <- function(df) {
  group_stats_func <- df %>%
    group_by(Group) %>%
    summarise(
      n = n(),
      goddess_n = sum(info_gender.factor == "female", na.rm = TRUE),
      goddess = sum(info_gender.factor == "female", na.rm = TRUE)/n())
  
  overall_stats_func <- df %>%
    summarise(
      group = NULL,
      n = n(),
      goddess_n = sum(info_gender.factor == "female", na.rm = TRUE),
      goddess = sum(info_gender.factor == "female", na.rm = TRUE)/n())
  
  return(bind_rows(group_stats_func, overall_stats_func))
  
}

stats_ethnicity(metadata_pp)
stats_ethnicity(metadata_do)

chisq.test(metadata_pp$Group, metadata_pp$info_gender.factor)
chisq.test(metadata_do$Group, metadata_do$info_gender.factor)

```


## Disease duration
```{r duration Stats, echo = TRUE}
stats_func <- function(df) {
  group_stats_func <- df %>%
    group_by(Group) %>%
    summarise(
      n = n(),
      median = round(median(Disease_Duration_.months., na.rm = TRUE)),
      IQR = round(IQR(Disease_Duration_.months., na.rm = TRUE))
    )
  
  overall_stats_func <- df %>%
    summarise(
      group = NULL,
      n = n(),
      median = round(median(Disease_Duration_.months., na.rm = TRUE)),
      IQR = round(IQR(Disease_Duration_.months., na.rm = TRUE))
    )
  
  return(bind_rows(group_stats_func, overall_stats_func))
  
}

stats_func(metadata_pp)
stats_func(metadata_do)
kruskal.test(Disease_Duration_.months. ~ Group, data = metadata_pp)
kruskal.test(Disease_Duration_.months. ~ Group, data = metadata_do)

```


## Diagnosis paste
```{r Diagnosis Stats, echo = TRUE}
stats_func <- function(df) {
  group_stats_func <- df %>%
    group_by(Group) %>%
    summarise(
      n = n(),
      median = round(median(Diagnosis_Duration_.months., na.rm = TRUE)),
      IQR = round(IQR(Diagnosis_Duration_.months., na.rm = TRUE))
    )
  
  overall_stats_func <- df %>%
    summarise(
      group = NULL,
      n = n(),
      median = round(median(Diagnosis_Duration_.months., na.rm = TRUE)),
      IQR = round(IQR(Diagnosis_Duration_.months., na.rm = TRUE))
    )
  
  return(bind_rows(group_stats_func, overall_stats_func))
  
}

stats_func(metadata_pp)
stats_func(metadata_do)
kruskal.test(Diagnosis_Duration_.months. ~ Group, data = metadata_pp)
kruskal.test(Diagnosis_Duration_.months. ~ Group, data = metadata_do)

```


## EDSS paste
```{r EDSS Stats, echo = TRUE}
stats_func <- function(df) {
  group_stats_func <- df %>%
    group_by(Group) %>%
    summarise(
      n = n(),
      median = round(median(edss.v0, na.rm = TRUE)),
      IQR = round(IQR(edss.v0, na.rm = TRUE))
    )
  
  overall_stats_func <- df %>%
    summarise(
      group = NULL,
      n = n(),
      median = round(median(edss.v0, na.rm = TRUE)),
      IQR = round(IQR(edss.v0, na.rm = TRUE))
    )
  
  return(bind_rows(group_stats_func, overall_stats_func))
  
}

stats_func(metadata_pp)
stats_func(metadata_do)
kruskal.test(edss.v0 ~ Group, data = metadata_pp)
kruskal.test(edss.v0 ~ Group, data = metadata_do)

```


## T2 lesions
```{r EDSS Stats, echo = TRUE}
stats_func <- function(df) {
  group_stats_func <- df %>%
    group_by(Group) %>%
    summarise(
      n = n(),
      median = round(median(mri_t2call.v0, na.rm = TRUE)),
      IQR = round(IQR(mri_t2call.v0, na.rm = TRUE))
    )
  
  overall_stats_func <- df %>%
    summarise(
      group = NULL,
      n = n(),
      median = round(median(mri_t2call.v0, na.rm = TRUE)),
      IQR = round(IQR(mri_t2call.v0, na.rm = TRUE))
    )
  
  return(bind_rows(group_stats_func, overall_stats_func))
  
}

stats_func(metadata_pp)
stats_func(metadata_do)

kruskal.test(mri_t2call.v0 ~ Group, data = metadata_pp)
kruskal.test(mri_t2call.v0 ~ Group, data = metadata_do)

```



## BMI paste
```{r BMI Stats, echo = TRUE}
stats_func <- function(df) {
  group_stats_func <- df %>%
    group_by(Group) %>%
    summarise(
      n = n(),
      median = round(median(bmi.v0, na.rm = TRUE), 1),
      IQR = round(IQR(bmi.v0, na.rm = TRUE), 1)
    )
  
  overall_stats_func <- df %>%
    summarise(
      group = NULL,
      n = n(),
      median = round(median(bmi.v0, na.rm = TRUE), 1),
      IQR = round(IQR(bmi.v0, na.rm = TRUE), 1)
    )
  
  return(bind_rows(group_stats_func, overall_stats_func))
  
}

stats_func(metadata_pp)
stats_func(metadata_do)

kruskal.test(bmi.v0 ~ Group, data = metadata_pp)
kruskal.test(bmi.v0 ~ Group, data = metadata_do)
```


## VitaminD
```{r VitaminD Stats, echo = TRUE}
stats_func <- function(df) {
  group_stats_func <- df %>%
    group_by(Group) %>%
    summarise(
      n = n(),
      mean = round(mean(lab_vitamind.v0, na.rm = TRUE), 1),
      sd = round(sd(lab_vitamind.v0, na.rm = TRUE), 1)
    )
  
  overall_stats_func <- df %>%
    summarise(
      group = NULL,
      n = n(),
      mean = round(mean(lab_vitamind.v0, na.rm = TRUE), 1),
      sd = round(sd(lab_vitamind.v0, na.rm = TRUE), 1)
    )
  
  return(bind_rows(group_stats_func, overall_stats_func))
  
}

stats_func(metadata_pp)
stats_func(metadata_do)
kruskal.test(lab_vitamind.v0 ~ Group, data = metadata_pp)
kruskal.test(lab_vitamind.v0 ~ Group, data = metadata_do)
```


## Fatigue paste
```{r Fatigue Stats, echo = TRUE}
stats_func <- function(df) {
  group_stats_func <- df %>%
    group_by(Group) %>%
    summarise(
      n = n(),
      median = round(median(fss_score.v0, na.rm = TRUE), 1),
      IQR = round(IQR(fss_score.v0, na.rm = TRUE), 1)
    )
  
  overall_stats_func <- df %>%
    summarise(
      group = NULL,
      n = n(),
      median = round(median(fss_score.v0, na.rm = TRUE), 1),
      IQR = round(IQR(fss_score.v0, na.rm = TRUE), 1)
    )
  
  return(bind_rows(group_stats_func, overall_stats_func))
  
}

stats_func(metadata_pp)
stats_func(metadata_do)
kruskal.test(fss_score.v0 ~ Group, data = metadata_pp)
kruskal.test(fss_score.v0 ~ Group, data = metadata_do)
```


## BDI paste
```{r BMI Stats, echo = TRUE}
stats_func <- function(df) {
  group_stats_func <- df %>%
    group_by(Group) %>%
    summarise(
      n = n(),
      median = round(median(bdi2_score_corrected.v0, na.rm = TRUE), 1),
      IQR = round(IQR(bdi2_score_corrected.v0, na.rm = TRUE), 1)
    )
  
  overall_stats_func <- df %>%
    summarise(
      group = NULL,
      n = n(),
      median = round(median(bdi2_score_corrected.v0, na.rm = TRUE), 1),
      IQR = round(IQR(bdi2_score_corrected.v0, na.rm = TRUE), 1)
    )
  
  return(bind_rows(group_stats_func, overall_stats_func))
  
}

stats_func(metadata_pp)
stats_func(metadata_do)
kruskal.test(bdi2_score_corrected.v0 ~ Group, data = metadata_pp)
kruskal.test(bdi2_score_corrected.v0 ~ Group, data = metadata_do)
```




## SDMT paste
```{r BMI Stats, echo = TRUE}
stats_func <- function(df) {
  group_stats_func <- df %>%
    group_by(Group) %>%
    summarise(
      n = n(),
      median = round(median(sdmt_result.v0, na.rm = TRUE), 1),
      IQR = round(IQR(sdmt_result.v0, na.rm = TRUE), 1)
    )
  
  overall_stats_func <- df %>%
    summarise(
      group = NULL,
      n = n(),
      median = round(median(sdmt_result.v0, na.rm = TRUE), 1),
      IQR = round(IQR(sdmt_result.v0, na.rm = TRUE), 1)
    )
  
  return(bind_rows(group_stats_func, overall_stats_func))
  
}

stats_func(metadata_pp)
stats_func(metadata_do)
kruskal.test(sdmt_result.v0 ~ Group, data = metadata_pp)
kruskal.test(sdmt_result.v0 ~ Group, data = metadata_do)
```



## DMT
```{r DMT Stats, echo = TRUE}
stats_func <- function(df) {
  group_stats_func <- df %>%
    group_by(Group) %>%
    summarise(
      n = n(),
      goddess_n = sum(No_DMT == T),
      goddess = sum(No_DMT == T)/n()*100)
  
  overall_stats_func <- df %>%
    summarise(
      group = NULL,
      n = n(),
      goddess_n = sum(No_DMT == T),
      goddess = sum(No_DMT == T)/n()*100)
  
  return(bind_rows(group_stats_func, overall_stats_func))
  
}

stats_func(metadata_pp)
stats_func(metadata_do)
chisq.test(metadata_pp$Group, metadata_pp$No_DMT)
chisq.test(metadata_do$Group, metadata_do$No_DMT)
```


## Single Meds
```{r Meds Stats, echo = TRUE}
stats_func <- function(df, med_name) {
  group_stats_func <- df %>%
    group_by(Group) %>%
    summarise(
      n = n(),
      sum = sum({{med_name}} == T),
      percent = round(sum({{med_name}} == T)/n()*100))
  
  overall_stats_func <- df %>%
    summarise(
      group = NULL,
      n = n(),
      sum = sum({{med_name}} == T),
      percent = round(sum({{med_name}} == T)/n()*100))
  
  return(bind_rows(group_stats_func, overall_stats_func))
  
}
names(ms_meds_filtered)
stats_func(metadata_pp, No_DMT)
stats_func(metadata_do, Tecfidera)

chisq.test(metadata_pp$Group, metadata_pp$Aubagio)
chisq.test(metadata_do$Group, metadata_do$Rituximab)
```
